### 2010-09-08 paper/outpaper.tex

\documentclass[codesnippet]{jss}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% declarations for jss.cls %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% almost as usual
\author{Taylor B. Arnold\\Yale University \And 
        John W. Emerson\\Yale University}
\title{Non-parametric Goodness of Fit Tests for Discrete Distributions}

%% for pretty printing and a nice hypersummary also set:
\Plainauthor{Taylor Arnold, John Emerson} %% comma-separated
\Plaintitle{Non-parametric Goodness of Fit Tests for Discrete Distributions} %% without formatting
\Shorttitle{Non-parametric Goodness of Fit Tests for Discrete Distributions} %% a short title (if necessary)

%% an abstract and keywords
\Abstract{
  Here should be the abstract.
}
\Keywords{keywords, comma-separated, not capitalized, \proglang{Java}}
\Plainkeywords{keywords, comma-separated, not capitalized, Java} %% without formatting
%% at least one keyword must be supplied

%% publication information
%% NOTE: Typically, this can be left commented and will be filled out by the technical editor
%% \Volume{13}
%% \Issue{9}
%% \Month{September}
%% \Year{2004}
%% \Submitdate{2004-09-29}
%% \Acceptdate{2004-09-29}

%% The address of (at least) one author should be given
%% in the following format:
\Address{
  Taylor B. Arnold\\
  24 Hillhouse Ave.\\
  New Haven, CT 06511, USA\\
  E-mail: \email{taylor.arnold@yale.edu}\\
  URL: \url{http://euler.stat.yale.edu/~tba3}
}
%% It is also possible to add a telephone and fax number
%% before the e-mail in the following format:
%% Telephone: +43/1/31336-5053
%% Fax: +43/1/31336-734

%% for those who use Sweave please include the following line (with % symbols):
%% need no \usepackage{Sweave.sty}

%% end of declarations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%% include your article here, just as usual
%% Note that you should use the \pkg{}, \proglang{} and \code{} commands.

\section[About Java]{About \proglang{Java}}
%% Note: If there is markup in \(sub)section, then it has to be escape as above.

\end{document}